---
author: "Jack Fu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{recountNNLS}
  %\usepackage[UTF-8]{inputenc}
---


## Overview

This package provides an interface to estimate transcript abundances of any samples quantified by the aligner *rail*. This method is a non-negative least squares (NNLS) estimation that calculates the number of reads that originated from each transcript of the coding portion of the GencodeV25 transcriptome. The model does not require raw aligned `bam` files, but is content with compressed coverage statistics primarily stored in `bigwig` formats.

All of the SRA projects compiled by *recount2* already have transcript expression pre-computed through this method and is accessible through *recount2*. To replicate the abundance estimation of any SRA project in *recount2*, the user only needs to supply the SRA project id. Otherwise, users can supply the necessary information as outlined further below to utilize this package to carry out transcript abundance estimation.

## Quantifying Samples on recount2

To use our NNLS model on the samples on recount2, all one needs is the SRA project id.

```{r, eval=F}
library(recountNNLS)

## Specify a SRA project and download the relevant path data from recount2
project = 'SRP063581'
pheno = processPheno(project)

## Main NNLS workhorse function to create a RSE of transcript abundance
rse_tx = recountNNLS(pheno)

```
## Data Not Yet Part of recount2

If not quantifying the transcript expression of a project already compiled by *recount2*, the user will need to provide:

1. A manifest of 5 columns is required. The columns are: project name (project); sample id (run); path to bigwig file resulting from *rail* alignment (bigwit_path); *rls* the read length of the sample; and whether the sample is paired end (paired_end).
2. A path to the junction table created by *rail*. It is usually located in the folder `cross_sample_results/junctions.tsv.gz`.

```{r, echo =F}
base='/dcl01/leek/data/ta_poc/geuvadis/simulation/37_1/rail/rail-rna_out/'
manifest = data.frame(project = "test", run = c("sample_01", "sample_02", "sample_03", "sample_04"), 
                      bigwig_path = c('coverage_bigwigs/sample_01.bw', 'coverage_bigwigs/sample_02.bw', 
                               'coverage_bigwigs/sample_03.bw', 'coverage_bigwigs/sample_04.bw'),
                      rls = c(37, 37, 37, 37), paired_end=F)
junction_path =  'cross_sample_results/junctions.tsv.gz'
```
```{r}
manifest
junction_path
```

To obtain the necessary inputs for the linear model, one would apply:
```{r, eval=F}
library(recountNNLS)
## Analyze the manifest input to format for downstream use
pheno = processPheno(manifest)

## Main NNLS workhorse function to create a RSE of transcript abundance
rse_tx = recountNNLS(pheno, jx_file=junction_path, cores=1)
```

## Deliverables

The output of `recountNNLS()` is a `RangeSummarizedExperiment` where:

* Each row corresponds to a transcript
* Each column corresponds to a run (sample)
* `rowRanges()` can be used to access a `GRangesList` annotation of the transcripts
* `assays()` returns a list of length 2 that can be access with `$`
      + `$counts` returns a matrix of estimated transcript abundance on a number of reads scale
      + `$se` returns a matrix of estimated standard error of the abundance



