---
author: "Jack Fu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Title of your vignette}
  %\usepackage[UTF-8]{inputenc}
---

## Overview

This pacakge provides the code for implementing the non-negative least squares (NNLS) estimation of transcript abundance based on the reduced-representation output from *rail*. All of the SRA projects quantified by *rail* and compiled by the *recount2* project can be directly quantified by inputing just the *Sequencing Read Archive* (SRA) project name. Otherwise, users can supply a manifest that contains the necessary paths to the bigwig and junction tables that *rail* outputs.

## Manifest

If not quantifying the transcript expression of a project already compiled by *recount2*, a manifest of 3 columns is required. The first column will list the names of the samples as well as an etnry for the junctions *rail* detected. The second column will list the path to the bigwig files of each sample as well as the junction file. The third column will be the sequencing read length of each sample.

```{r, echo =F}
base='/dcl01/leek/data/ta_poc/geuvadis/simulation/37_1/rail/rail-rna_out/'
manifest = data.frame(run = c("sample_01", "sample_02", "sample_03", "sample_04"), 
                      bigwig_path = c('coverage_bigwigs/sample_01.bw', 'coverage_bigwigs/sample_02.bw', 
                               'coverage_bigwigs/sample_03.bw', 'coverage_bigwigs/sample_04.bw'),
                      avg_read_length = c(37, 37, 37, 37))
junction_path =  'cross_sample_results/junctions.tsv.gz'
```
```{r}
manifest
junction_path
```





